### 2024-04-23 parts/conclusion.tex
mise en forme
almost to 12 pages

Let us conclude with a more general remark on the status of commutative input
in the study of unary output polyregular functions. Notice that all semantic
characterizations of subclasses of $\ZPoly$ (\kl{combinatorial}, \kl{ultimately
polynomial}, pumpable) arise from pre-composition with a \kl{commutative}
\kl{star-free polyregular function}, as formally stated in
\cref{pre-compose-growth-commut:lemma,pre-compose-sf-commut:lemma}. Following
this idea, we conjecture that $\ZPoly$ and $\NPoly$ can be described as
combinations of commutative functions, leading to
\cref{z-poly-commutative-encoding:conj}. This conjecture would play the role of
\cref{decompose-polynomial:lem} in the study of functions in $\ZPoly$.

\begin{lemma}
    \label{pre-compose-growth-commut:lemma}
    Let $f \in \ZRat$, and $k \in \Nat$. Then,
    $f \in \ZPoly[k]$ if and only if 
    for every \kl{commutative} \kl{star-free polyregular function} $h$
            of \kl{growth rate} $l \in \Nat$,
            $(f \circ h) \in \ZPoly[k+l-1]$.
\end{lemma}

\begin{lemma}
    \label{pre-compose-sf-commut:lemma}
    Let $f \in \ZPoly$. Then, $f \in \ZSF$,
    if and only if for every \kl{commutative} \kl{star-free polyregular function} $h$,
            $(f \circ h) \in \ZSF$.
\end{lemma}

@@

In the specific subclass of \emph{polyblind functions}
\cite{LENP21,DOUE22}, parts of the conjecture holds,
because they essentially behave as commutative functions \cite[Theorem
6.12]{DOUE23}. Indeed, a suitable application of Parikh's Theorem
\cite{PARI66} guarantees the existence of the functions $\mathsf{enc}$
and $\mathsf{dec}$, although they may fail to be  \kl{polyregular functions}.

\begin{lemma}
    \label{z-poly-commutative-encoding:remark}
    The \cref{z-poly-commutative-encoding:conj} holds for the
    subclass of polyblind functions when dropping the assumption that
    $\mathsf{enc}$ and $\mathsf{dec}$ are \kl{polyregular functions}
    (resp. \kl{star-free polyregular functions}).
\end{lemma}

>>>>

Let us conclude with a more general remark on the status of commutative input
in the study of unary output polyregular functions. Notice that semantic
characterizations of subclasses of $\ZPoly$ (\kl{combinatorial}, \kl{ultimately
polynomial}) arise from pre-composition with a \kl{commutative}
\kl{star-free polyregular function}, as formally stated in
\cref{pre-compose-growth-commut:lemma,pre-compose-sf-commut:lemma}. Following
this idea, we conjecture that $\ZPoly$ and $\NPoly$ can be described as
combinations of commutative functions, leading to
\cref{z-poly-commutative-encoding:conj}, which would be a non-commutative
version of \cref{decompose-polynomial:lem}. For the subclass of 
\emph{polyblind functions} \cite{LENP21,DOUE22}, parts of the conjecture holds,
because they essentially behave as commutative functions \cite[Theorem
6.12]{DOUE23}.

%Indeed, a suitable application of Parikh's Theorem \cite{PARI66}
%guarantees the existence of the functions $\mathsf{enc}$ and $\mathsf{dec}$,
%although they may fail to be  \kl{polyregular functions}.

\begin{lemma}
    \label{pre-compose-growth-commut:lemma}
    Let $f \in \ZRat$, and $k \in \Nat$. Then,
    $f \in \ZPoly[k]$ if and only if 
    for every \kl{commutative} \kl{star-free polyregular function} $h$
            of \kl{growth rate} $l \in \Nat$,
            $(f \circ h) \in \ZPoly[k+l-1]$.
    \proofref{pre-compose-growth-commut:lemma}
\end{lemma}

\begin{lemma}
    \label{pre-compose-sf-commut:lemma}
    Let $f \in \ZPoly$. Then, $f \in \ZSF$,
    if and only if for every \kl{commutative} \kl{star-free polyregular function} $h$,
            $(f \circ h) \in \ZSF$.
    \proofref{pre-compose-sf-commut:lemma}
\end{lemma}

@@

\begin{lemma}
    \label{z-poly-commutative-encoding:remark}
    The \cref{z-poly-commutative-encoding:conj} holds for the
    subclass of polyblind functions when dropping the assumption that
    $\mathsf{enc}$ and $\mathsf{dec}$ are \kl{polyregular functions}
    (resp. \kl{star-free polyregular functions}).
    \proofref{z-poly-commutative-encoding:remark}
\end{lemma}

### 2024-04-23 parts/correction.tex
almost to 12 pages

Let us now illustrate how the \kl{discrete differentiation} operator interacts
with \kl{maximal monomials} with respect to the \kl{Hoare ordering}: it
extracts \emph{sub-maximal} monomials from a given polynomial.

\begin{lemma}
    \label{derivation-simplifies:lemma}
    For all $P \in \Rel[\vec{X}]$ that are non-constant,
    for all $K \in \Nat$,
    $\MaximalMonomials(\Diff{K}{P}) \hoarele
    \MaximalMonomials(P)$.
    \proofref{derivation-simplifies:lemma}
\end{lemma}

The following \cref{derivation-stabilises-correct:lem} is the main
combinatorial argument of this section. It leverages the positivity of the
\kl{maximal monomials} to compute a shift large enough to make lower degree
coefficients positive. This lemma will be lifted to a given polynomial $P$ by
noticing that if $K \in \Nat$, and $P = P_1 + P_2$, then $\translate{K}(P) =
P_1 + \Diff{K}{P_1} + \translate{K}(P_2)$.

@@

%In the actual statement of \cref{corrected-version:thm}, the representation of
%polynomials in $\CorrectPoly$ is strengthened from \kl{$\Nat$-polyregular
%functions} to \kl{star-free $\Nat$-polyregular functions}. This is not
%surprising, given the conjecture that \kl{ultimately polynomial}
%\kl{$\Nat$-polyregular functions} (the formal definition will be given in
%\cref{star-free:sec}) were conjectured to be star free
%(\cref{zsf-nsf:conjecture}), and that polynomials are always \kl{ultimately
%polynomial}.

\begin{lemma}
    \label{lem:correct-to-n-rat}
    Let $\vec{X}$ be a tuple of indeterminates,
    and let $P \in \Rel[\vec{X}]$.
    If $P \in \CorrectPoly$, then $P$ is \kl{represented}
    by a \kl{star-free $\Nat$-polyregular function},
    which can be explicitly constructed from $P$.
\end{lemma}

While \cref{lem:correct-to-n-rat} provides effective conversion, it does not
explicitly state that the membership is decidable to keep the proof clearer. A
similar proof scheme can be followed to conclude that membership is decidable,
and even show that elements in $\CorrectPoly$ are, up to suitable translations,
polynomials in $\Nat[\vec{X}]$. Beware that partial applications are still
needed in this characterization, as \cref{bad-poly-translate:ex} illustrates.

@@

We are now ready to state the corrected and generalized version of
\cref{karh:thm}, which is the main technical contribution of the paper.

@@

The counter example of \cref{thm:counter-example} used three indeterminates,
and this is not a coincidence: in the particular case of at most two
indeterminates, the classes $\CorrectPoly$ and $\CoveredPoly$ coincide. In
particular, the examples appearing in \cite{KARH77} are not invalidated, as
they all use at most two indeterminates. This is because, in the
univariate case, being non-negative and having non-negative maximal coefficient
implies being an \kl{$\Nat$-rational polynomial}.

>>>>

The key lemma of this section is \cref{lem:correct-to-n-rat}, which is proved
by induction on the \kl{maximal monomials} of a polynomial $P$. The core
combinatorial argument allowing to perform the induction is
\cref{derivation-stabilises-correct:lem}, which is proved by a straightforward
binomial expansion. Using this lemma, the induction is performed as follows:
after decomposing a polynomial $P$ into a sum of \kl{maximal monomials} $P_1$
and the rest $P_2$, notice that for all $K \in \Nat$, $\translate{K}(P) = P_1 +
\Diff{K}{P_1} + \translate{K}(P_2)$. Then, because $P \in \CorrectPoly$, $P_1$
is a \kl{$\Nat$-rational polynomial}, furthermore, we can use the induction
hypothesis on $\Diff{K}{P_1} + \translate{k}(P_2)$ since it has lower degree
monomials and is in $\CorrectPoly$, finally we can also apply the induction
hypothesis to $\restr{P}{\nu}$ where $\nu$ is a partial function from $\vec{X}$
to $\set{0, \dots, K}$ fixing at least one variable. It is then just a matter
of combining these functions to obtain a \kl{$\Nat$-polyregular function} that
\kl{represents} $P$. Notice that in \cref{lem:correct-to-n-rat}, polynomials
can be represented using \kl{star-free $\Nat$-polyregular functions}, which is
a stronger statement than the one in \cref{karh:thm}, but is in line with the
idea that polynomials are not periodic functions (for instance, in
\cref{zsf-nsf:conjecture}).

@@

\begin{lemma}
    \label{lem:correct-to-n-rat}
    Let $\vec{X}$ be a tuple of indeterminates,
    and let $P \in \Rel[\vec{X}]$.
    If $P \in \CorrectPoly$, then $P$ is \kl{represented}
    by a \kl{star-free $\Nat$-polyregular function},
    which is computable given $P$.
\end{lemma}

While \cref{lem:correct-to-n-rat} provides effective conversion, it does not
explicitly state that the membership is decidable to keep the proof clearer. A
similar proof scheme can be followed to conclude that membership is decidable,
and even show that elements in $\CorrectPoly$ are, up to suitable translations,
polynomials in $\Nat[\vec{X}]$. Beware that partial applications are still
needed in this characterization, as \cref{bad-poly-translate:ex} illustrates.
Effectiveness was the last missing piece of our main result
\cref{corrected-version:thm}, which corrects and generalizes \cref{karh:thm}.

@@

@@

For completeness, let us remark that the counter example of
\cref{thm:counter-example} uses three indeterminates, and this is not a
coincidence: in the particular case of at most two indeterminates, the classes
$\CorrectPoly$ and $\CoveredPoly$ coincide. In particular, the examples
appearing in \cite{KARH77} are not invalidated, as they all use at most two
indeterminates. This is because, in the univariate case, being non-negative and
having non-negative maximal coefficient implies being an \kl{$\Nat$-rational
polynomial}.

### 2024-04-23 parts/counter.tex
almost to 12 pages

\AP A polynomial $P \in \Rel[X_1, \dots, X_n]$ is an \intro{$\Nat$-rational
polynomial} if it is \kl{represented} by a \kl{$\Nat$-polyregular function}.
The analogue notion of \intro{$\Rel$-rational polynomials} is not of particular
interest, since every polynomial $P$ is \kl{represented} by a
\kl{$\Rel$-polyregular function}. It is an easy check that polynomials with
coefficients in $\Nat$ are \kl{$\Nat$-rational polynomials}
(\cref{n-poly-n-poly:example}). However, \cref{negative-but-npoly:ex} provides
a polynomial with negative coefficients that is an \kl{$\Nat$-rational
polynomial}. As a consequence, the class of interest strictly contains
$\Nat[\vec{X}]$.

@@

The goal of this section is to provide semantic characterizations of
\kl{$\Nat$-rational polynomials}, decide if a polynomial $P$ is an
\kl{$\Nat$-rational polynomial}, and if so, explicitly compute an
\kl{$\Nat$-polyregular function} $f$ that \kl{represents} $P$. To that end, let
us now introduce the characterization of \kl{$\Nat$-rational polynomials} given
by \cite{KARH77}, which is stated using a class $\CoveredPoly[\vec{X}]$ of
polynomials with multiple indeterminates.

>>>>

\AP Before deciding $\NPoly$ inside $\ZPoly$ (which will be done in
\cref{beyond-polynomials:sec}), let us restrict our attention to simpler
functions, namely polynomials. A polynomial $P \in \Rel[X_1, \dots, X_n]$ is an
\intro{$\Nat$-rational polynomial} if it is \kl{represented} by a
\kl{$\Nat$-polyregular function}. It is an easy check that polynomials with
coefficients in $\Nat$ are \kl{$\Nat$-rational polynomials}
(\cref{n-poly-n-poly:example}). However, \cref{negative-but-npoly:ex} provides
a polynomial with negative coefficients that is an \kl{$\Nat$-rational
polynomial}.

The analogue notion of \intro{$\Rel$-rational
polynomials} is not of particular interest, since every polynomial $P$ is
\kl{represented} by a \kl{$\Rel$-polyregular function}. Our goal is to decide
if a \kl{$\Rel$-rational polynomial} (so, any polynomial) is an
\kl{$\Nat$-rational polynomial}, and if so, to compute an
\kl{$\Nat$-polyregular function} that \kl{represents} it. 
This particular problem was claimed to be solved in  \cite{KARH77}, 
using a characterization restated in \cref{karh:thm}.

@@

### 2024-04-23 parts/starfree.tex
almost to 12 pages

\begin{definition}[Ultimately polynomial]
    \label{ultimately-polynomial:def}
    Let $\Sigma$ be a finite alphabet. 
    A function $f \colon \Sigma^* \to \Rel$
    is \intro{ultimately polynomial}
    whenever there exists $N_0 \in \Nat$ such that
    for all $n \in \Nat$
    and for all words $\alpha_0, w_1, \alpha_1, \cdots, \alpha_{n-1}, w_n, \alpha_n
    \in \Sigma^*$, there exists a polynomial $P \in \Rel[X_1, \dots, X_n]$
    such that
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^{n} w_i^{X_i} \alpha_i
        \right)
        = 
        P(X_1, \dots, X_n)
        \quad 
        \forall X_1, \dots, X_n \geq N_0
        \quad .
    \end{equation*}
\end{definition}

>>>>

\begin{definition}[Ultimately polynomial]
    \label{ultimately-polynomial:def}
    Let $\Sigma$ be a finite alphabet. 
    A function $f \colon \Sigma^* \to \Rel$
    is \intro{ultimately polynomial}
    when there exists $N_0 \in \Nat$ such that
    for all $n \in \Nat$
    and for all words $\alpha_0, w_1, \alpha_1, \cdots, \alpha_{n-1}, w_n, \alpha_n
    \in \Sigma^*$, there exists a polynomial $P \in \Rel[X_1, \dots, X_n]$
    such that
    \begin{equation*}
        f\left(
            \alpha_0 \prod_{i = 1}^{n} w_i^{X_i} \alpha_i
        \right)
        = 
        P(X_1, \dots, X_n)
        \quad 
        \forall X_1, \dots, X_n \geq N_0
        \quad .
    \end{equation*}
\end{definition}

### 2024-04-23 parts/transducers.tex
almost to 12 pages

\AP The goal of this section is to provide tools that do not require the
assumption of \kl{commutativity}. To that end, we introduce the notion of
\kl{$k$-residual transducer}, that generalizes the notion of residual
transducer introduced for \kl{$\Rel$-polyregular functions} in \cite{CDTL23}.
This notion of residual transducer has its own interest, as it shifts the
attention from (compatible) equivalence relations of finite index that are the
classical tool in automata theory, towards (compatible) order relations that
are \kl{well-quasi-ordered}, which is the order counterpart of having finite
index. Let us recall that a sequence $\seqof{u_i}{i \in \Nat}$ of elements in a
quasi-ordered set $(X, \leq)$ is \intro{good} whenever there exist $i < j$ such
that $u_i \leq u_j$. The set $X$ is a \intro{well-quasi-ordering} when every
infinite sequence is \reintro{good}. A sequence is \intro{bad} when it is not
\reintro{good}. The notion of \reintro{good sequences} can also be applied to
binary relations that are not orderings, and a binary relation $R$ for which
every infinite sequence is \reintro{good} is said to be \intro{well}
\cite{MELL98}.

@@

\AP Given a function $f$, our goal is to leverage $\resleq{f}{k}$ to build a
canonical \kl{$\NPoly[k-1]$-transducer} that \kl{computes} $f$. The naïve
approach is to consider as states some representatives for the $\resleq{f}{k}$
minimal elements of $\Sigma^*$ (assuming there are finitely many of them),
%all computing the same function (\cref{good-residual-ordering:fact}),
and define transitions by letting $\delta(u, a)$ be \emph{one} state $v$ such
that $v \resleq{f}{k} ua$. This naïve approach does not yield a canonical
model, as illustrated by the two distinct \kl{$\NPoly[0]$-transducers} of
\cref{non-canonical-transd:fig} computing the function $\BadExOk$ of
\cref{non-canonical-transd:ex}, and having as states minimal elements for
$\resleq{\BadExOk}{0}$. To obtain a canonical notion, we enforce that the set
of states is a \intro{downwards closed} subset of $\Sigma^*$ for the prefix
ordering $(\prefleq)$, that is, every prefix of a state is also a state.

@@

Because of the restriction on prefixes, the \kl{$k$-residual transducer} is
actually unique when it exists. We now introduce \cref{residual:algo} to
compute the \kl{$k$-residual transducer} given a function $f$. Notice that this
algorithm requires the ability to test if a function belongs to $\NPoly[k]$,
which is only known to be feasible for \kl{commutative} \kl{polyregular
functions}. However, the termination of this algorithm also proves the
existence of the \kl{$k$-residual transducer}. The key argument proving the
termination of \cref{residual:algo} is based on the fact that for a function $f
\in \NPoly[k]$, the quasi-ordering $(\resleq{f}{k})$ is a
\kl{well-quasi-ordering}.

@@

\begin{theorem}
    \label{non-commutative-npoly:thm}
    Let $f \in \ZPoly$ be a non-negative function, 
    and $k \in \Nat$,
    the following are equivalent:
    \begin{enumerate}
        \item \label{n-poly-1-transd:item} $f$ is \kl{computed}
            by a \kl{$\NPoly[k-1]$-transducer};
        \item \label{n-poly-k:item} $f \in \NPoly[k]$;
        \item \label{n-poly-wqo:item} $(\Sigma^*, \resleq{f}{k})$ is a
            \kl{well-quasi-ordering};
        \item \label{n-poly-well:item} every $\prefleq$-increasing sequence
            of $\Sigma^*$  is a \kl{good sequence}
            for $\resleq{f}{k}$;
        \item \label{n-poly-residual:item} $f$ is \kl{computed} by a
            \kl{$k$-residual transducer}.
    \end{enumerate}
    If $f$ is assumed to be \kl{commutative}, the  
    properties are decidable, and the conversions effective.
    \proofref{non-commutative-npoly:thm}
\end{theorem}

\AP Unfortunately, the \kl{residual transducer} does not seem to provide an
algorithm to decide $\NSF$ inside $\NPoly$ in the general setting. While we can
prove that for functions in $\NPoly[0]$, aperiodicity is equivalent to the
absence of counters in the \kl{residual transducer} of the function (see
\cref{aperiodic-iff-residual:lem}), this result does not generalize to higher
growth rates, as witnessed by \cref{non-aperiodic-residual-transd:ex}.

@@

In the example, no \kl{$\NPoly[0]$-transducer} with the minimal number of
states computing the function is \kl{counter-free}, which suggests that a more
sophisticated object is needed to observe aperiodicity. In the spirit of the
characterization of $\NPoly$ based the $(\resleq{f}{k})$ quasi-ordering
(\cref{non-commutative-npoly:thm}), we introduce the notion of \intro{aperiodic
ordering} of $\Sigma^*$ as follows: $(\Sigma^*, \leq)$ is \reintro[aperiodic
ordering]{aperiodic} whenever for all $u, w \in \Sigma^*$, there exists $N_0
\in \Nat$, such that the sequence $\seqof{uw^n}{n \geq N_0}$ is increasing.
This leads to the following \cref{sf-no-periods-on-sequences:conj}, of which
the easy implication is already known to hold
(\cref{sf-no-periods-on-sequences:lemma}), and requires the introduction of the
\emph{star-free variant} $(\intro*\resleqsf{f}{k})$ of $(\resleq{f}{k})$,
defined by $u \reintro*\resleqsf{f}{k} v$ whenever $(\app{f}{v} - \app{f}{u})
\in \NSF[k-1]$.

\begin{lemma}
    \label{sf-no-periods-on-sequences:lemma}
    Let $k \in \Nat$, and $f \in \SF[k]$. Then,
    $(\Sigma^*, \resleqsf{f}{k})$ is an
    \kl[aperiodic ordering]{aperiodic} \kl{well-quasi-ordering}.
    \proofref{sf-no-periods-on-sequences:lemma}
\end{lemma}

\begin{conjecture}
    \label{sf-no-periods-on-sequences:conj}
    For all $k \in \Nat$ and $f \colon \Sigma^* \to \Rel$,
    $f \in \NSF[k]$ if and only if $(\Sigma^*, \resleqsf{f}{k})$ is an \kl[aperiodic ordering]{aperiodic}
    \kl{well-quasi-ordering}.
\end{conjecture}

>>>>

\AP The goal of this section is to introduce new tools that do not require the
assumption of \kl{commutativity}. To that end, we propose a notion of
\kl{$k$-residual transducer}, that generalizes the notion of residual
transducer introduced for \kl{$\Rel$-polyregular functions} in \cite{CDTL23}.
This is done by shifting the attention from (compatible) equivalence relations
of finite index that are the classical tool in automata theory, towards
(compatible) order relations that are \kl{well-quasi-ordered}, which is the
order counterpart of having finite index. Let us recall that a sequence
$\seqof{u_i}{i \in \Nat}$ of elements in a quasi-ordered set $(X, \leq)$ is
\intro{good} whenever there exist $i < j$ such that $u_i \leq u_j$. The set $X$
is a \intro{well-quasi-ordering} when every infinite sequence is
\reintro{good}. A sequence is \intro{bad} when it is not \reintro{good}. The
notion of \reintro{good sequences} can also be applied to binary relations that
are not orderings, and a binary relation $R$ for which every infinite sequence
is \reintro{good} is said to be \intro{well} \cite{MELL98}.

@@

\AP Given a function $f$, our goal is to leverage $\resleq{f}{k}$ to build a
canonical \kl{$\NPoly[k-1]$-transducer} that \kl{computes} $f$. The idea is to
consider as states the minimal elements of $\Sigma^*$ for $\resleq{f}{k}$, and
define transitions by letting $\delta(u, a)$ be some state $v$ such that $v
\resleq{f}{k} ua$. To produce a canonical model, this has to be done carefully,
as illustrated by the two distinct \kl{$\NPoly[0]$-transducers} of
\cref{non-canonical-transd:fig} computing the function $\BadExOk$ of
\cref{non-canonical-transd:ex}, and having as states minimal elements for
$\resleq{\BadExOk}{0}$. To ensure unicity, we ask that the set of states is a
\intro{downwards closed} subset of $\Sigma^*$ for the prefix ordering
$(\prefleq)$, i.e. that every prefix of a state is also a state.

@@

We now introduce \cref{residual:algo} to
compute the \kl{$k$-residual transducer} given a function $f$. Notice that this
algorithm requires the ability to test if a function belongs to $\NPoly[k]$,
which is only known to be feasible for \kl{commutative} \kl{polyregular
functions}. However, the termination of this algorithm also proves the
existence of the \kl{$k$-residual transducer}. The key argument proving the
termination of \cref{residual:algo} is based on the fact that for a function $f
\in \NPoly[k]$, the quasi-ordering $(\resleq{f}{k})$ is a
\kl{well-quasi-ordering}.

@@

\begin{theorem}
    \label{non-commutative-npoly:thm}
    Let $f \in \ZPoly$ be a non-negative function, 
    and $k \in \Nat$,
    the following are equivalent:
    \begin{enumerate}
        \item \label{n-poly-1-transd:item} $f$ is \kl{computed}
            by a \kl{$\NPoly[k-1]$-transducer};
        \item \label{n-poly-k:item} $f \in \NPoly[k]$;
        \item \label{n-poly-wqo:item} $(\Sigma^*, \resleq{f}{k})$ is a
            \kl{well-quasi-ordering};
        \item \label{n-poly-well:item} every $\prefleq$-increasing sequence
            of $\Sigma^*$  is a \kl{good sequence}
            for $\resleq{f}{k}$;
        \item \label{n-poly-residual:item} $f$ is \kl{computed} by a
            \kl{$k$-residual transducer}.
    \end{enumerate}
    If $f$ is \kl{commutative}, the  
    properties are decidable, and the conversions effective.
    \proofref{non-commutative-npoly:thm}
\end{theorem}

\AP Unfortunately, the \kl{residual transducer} does not seem to provide an
algorithm to decide $\NSF$ inside $\NPoly$ in the general setting. While we can
prove that for functions in $\NPoly[0]$, aperiodicity is equivalent to the
absence of counters in the \kl{residual transducer} of the function (see
\cref{aperiodic-iff-residual:lem}), this result does not generalize to higher
growth rates, as witnessed by \cref{non-aperiodic-residual-transd:ex}, which
suggests that a more sophisticated object is needed to observe aperiodicity. In
the spirit of the characterization of $\NPoly$ based the $(\resleq{f}{k})$
quasi-ordering (\cref{non-commutative-npoly:thm}), we introduce the notion of
\intro{aperiodic ordering} of $\Sigma^*$ as follows: $(\Sigma^*, \leq)$ is
\reintro[aperiodic ordering]{aperiodic} whenever for all $u, w \in \Sigma^*$,
there exists $N_0 \in \Nat$, such that the sequence $\seqof{uw^n}{n \geq N_0}$
is increasing. This leads to the following
\cref{sf-no-periods-on-sequences:conj}, of which the easy implication is
already known to hold (\cref{sf-no-periods-on-sequences:lemma}), and requires
the introduction of the \emph{star-free variant} $(\intro*\resleqsf{f}{k})$ of
$(\resleq{f}{k})$, defined by $u \reintro*\resleqsf{f}{k} v$ whenever
$(\app{f}{v} - \app{f}{u}) \in \NSF[k-1]$.

@@

\begin{lemma}
    \label{sf-no-periods-on-sequences:lemma}
    Let $k \in \Nat$, and $f \in \NPoly[k]$. If $f \in \NSF[k]$, then
    $(\Sigma^*, \resleqsf{f}{k})$ is an
    \kl[aperiodic ordering]{aperiodic} \kl{well-quasi-ordering}.
    \proofref{sf-no-periods-on-sequences:lemma}
\end{lemma}

\begin{conjecture}
    \label{sf-no-periods-on-sequences:conj}
    The implication of \cref{sf-no-periods-on-sequences:lemma} is actually an equivalence.
    %For all $k \in \Nat$ and $f \colon \Sigma^* \to \Rel$,
    %$f \in \NSF[k]$ if and only if $(\Sigma^*, \resleqsf{f}{k})$ is an \kl[aperiodic ordering]{aperiodic}
    %\kl{well-quasi-ordering}.
\end{conjecture}
